\subsection{Uncertainty and failure modes}
\label{sec:result-unc}



The comparisons so far show that the corrected \xcode improves the agreement with TCCON. However, the confidence of the agreement between OCO-2 and TCCON within the uncertainties remains unclear. For every station-day, we form the difference $D = \overline{X^{\mathrm{DE}}_{\mathrm{CO2}}} - X^{\mathrm{TCCON}}_{\mathrm{CO2}}$ and give it an uncertainty from both sides: the ensemble's own predictive $\sigma$ on the OCO-2 side (scaled by a factor of up to 1.27 near cloud, calibrated on held-out data) and the TCCON measurement, temporal-variability, harmonization, and collocation terms on the other (Appendix~D, Table~D4). Because each station-day can carry its own small true offset — different air masses, seasons, and site conditions — we summarize the 75 differences with a random-effects average (DerSimonian--Laird estimator; \citealp{dersimonian1986meta}), which returns two numbers: the mean offset $\mu$ common to all cases, and the real case-to-case spread $\tau$ that remains after the stated uncertainties are accounted for. We find $\mu = -0.32 \pm 0.09$~ppm and $\tau = 0.52$~ppm (Appendix~D, Fig.~D5). The offset is at the level of the TCCON reference scale
itself, so we do not read it as a property of the correction. The spread, in contrast, mostly reflects how loosely OCO-2 and TCCON are matched in space and time: repeating the analysis at a 50~km
collocation radius shrinks $\tau$ from 0.52 to 0.22~ppm while $\mu$ stays within its confidence interval ($-0.45 \pm 0.07$~ppm). Two consequences follow for users of the corrected product. First,
improvement is not the same as agreement: only 1 of the 75 station-days passes a formal equivalence test with TCCON at $\pm$0.5~ppm (two one-sided tests, TOST), and 13 still differ significantly. Second, the per-case error bars alone are too optimistic — the residuals are on average 2.4 times larger than they predict at 100~km (1.8 times at 50~km) — and become consistent only when the between-case spread $\tau$ is added; we therefore recommend carrying $\tau$ as an additional error term in any downstream use.

The reliability boundary of the correction itself is narrow and, more importantly, visible from the inside. The strata with the largest remaining error — bright land surfaces, high aerosol load, and the high-latitude station-day bias — are under-corrected rather than mis-corrected, as established in Sect.~\ref{sec:result-tccon_val}. At the footprint level, the ensemble's predictive $\sigma$ points in the
useful direction: over land, the strong near-cloud corrections carry large $\sigma$ and almost never make a footprint worse (worsened fraction 0.06 in the highest-$\sigma$ decile), while the rare worsenings concentrate at small $\sigma$, where the predicted correction is itself small (worsened fraction 0.46 in the lowest-$\sigma$ decile, with small amplitudes). A downstream user can therefore trust the large corrections most, and treat $\sigma$ as a per-footprint reliability indicator — with the explicit limit that it describes the correction's own confidence and cannot detect the reference-scale offset and common-mode biases above, which bound the absolute accuracy of the corrected product at the few-tenths-of-a-ppm level.
